Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-27 paper/paper.Rmd
add forecasts ready for scoring
move absolute error discussion and add additional across horizon ae figure #17

```{r packages, cache = FALSE}
library(data.table)
library(purrr)
library(scoringutils)
library(lubridate)
library(forcats)
library(ggplot2)
library(ggridges)
library(patchwork)
library(scales)
library(here)
```

@@

@@

relative_wis_by_horizon <- scores |>
  summarise_scores(by = c("horizon", "model")) |>
  calc_relative_score(cols = c("horizon")) |>
  DT(, ris := round(relative_interval_score, 2))

@@

Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased.
Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.

@@

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  plot_rwis_location
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
This is true across forecast horizons but the magnitude of the difference increases.
If we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
When we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS.
This implies a similarly skewed distribution for the relaive point forecasts from the surrogate model point forecasts compared to the ensemble forecast which again  matches that observed when evaluating the full predictive distributions of both forecasts.

>>>>

```{r packages, cache = FALSE}
library(data.table)
library(purrr)
library(scoringutils)
library(lubridate)
library(forcats)
library(ggplot2)
library(ggridges)
library(ggh4x)
library(patchwork)
library(scales)
library(here)
```

@@

# Load forecasts processed ready for scoring
forecasts_ready_for_scoring <- fread(
  here("data", "forecasts_ready_for_scoring.csv")
)

@@

relative_wis_by_horizon <- scores |>
  summarise_scores(by = c("horizon", "model")) |>
  calc_relative_score(cols = c("horizon")) |>
  DT(, ris := round(relative_interval_score, 2)) |>
  DT(, interval_score := NULL)
setnames(
  relative_wis_by_horizon, c("relative_interval_score"),
  c("Weighted interval score")
)

relative_ae_by_horizon <- scores |>
  summarise_scores(by = c("horizon", "model")) |>
  DT(, interval_score := ae_median) |>
  calc_relative_score(cols = c("horizon")) |>
  DT(, ris := round(relative_interval_score, 2)) |>
  DT(, interval_score := NULL)
setnames(
  relative_ae_by_horizon, c("relative_interval_score", "ris"),
  c("Absolute error of the median", "raem")
)

relative_summary <- relative_wis_by_horizon |>
  merge(
    relative_ae_by_horizon, by = c("horizon", "model")
  ) |>
  melt(
    measure.vars = c("Weighted interval score", "Absolute error of the median")
  )

@@

Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.

@@

If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS. As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  (
    plot_relative_summary(relative_summary) +
    plot_rwis_location +
    plot_layout(heights = c(1, 7), guides = "collect")
  )
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
This is true across forecast horizons but the magnitude of the difference increases.
